\section{Discussion}\label{sec:discussion}

% Summarize what we’ve learned
With this paper we aimed to find a common language to categorise different ways to compute model weights, describe the different choices that have to be made together with their implications, and generally highlight their advantages and disadvantages.
We differentiate between two general approaches: ensemble and individual performance weighting. While in the former weights are computed based on the performance of a linear combination of all models, in the latter, weights are computed directly from the performance of each individual model. 

As we showed by means of toy examples, a practical advantage of the ensemble performance weighting approach is that it implicitly accounts for the issue of inter-model dependencies since each weight represents the relative contribution of the corresponding model to the weighted average instead of representing its performance as such. Put differently, the mean weight from the ensemble performance approach depends not only on how closely the model simulates the observed data, but also on the other models in the ensemble: If there is a model $m_j$ in the ensemble that provides the same information as a model $m_i$, the average weight for both models will be reduced accordingly. Similar to defining model independence in terms of model output similarities, the “independence” part in the resulting mean weight vector, obtained from applying the ensemble performance weighting approach, does not directly depend on the observational data. While some consider it an advantage of the independence weights to not depend on observational data \citep[e.g.][]{brunnerReducedGlobalWarming2020}, others argue that the distance to observational data should have an influence on the independence weights \citep[e.g.][]{abramowitzESDReviewsModel2019}. In any case, what can be noted is that neither the individual performance weighting nor the ensemble performance weighting approach takes into account whether two models represent the relevant physical processes for the target output variable in the same way or differently. Even if models make similar predictions, they perhaps still should be considered independent if their predictions result from different representations of the underlying processes. 

We further considered two additional aspects under the ensemble performance weighting approach, namely the use of multiple diagnostics to measure performance as well as the influence of the choice of the prior distribution over weights.
% multiple diagnostics
When using multiple diagnostics, we showed that applying them jointly so that the weights are tuned to maximize the overall likelihood, i.e.~encompassing all diagnostics, yields better results in terms of the summed RMSE between the weighted average model (using the mean posterior weight vector) and the observed data than applying the diagnostics sequentially and then taking the average of the computed weight vectors. The choice of which diagnostics are most appropriate for measuring performance is a big question on its own \citep[e.g., see][]{glecklerPerformanceMetricsClimate2008} that we have not addressed here. 
Since, in the case of climate projections, we are assessing the performance of physical models, it makes sense to analyze variables that are related to the basic dynamic processes relevant for the target \citep{katzenbergerDevelopingGuidelinesWorking2025}. In any case, it should always be made clear that there is no universal answer and that any probabilistic projections are conditional on the metrics and diagnostic climate variables that were used. Furthermore, when multiple diagnostics are considered, these should be as independent of one another as possible since otherwise we would run into the same problem that we have with co-dependent models in the individual performance weighting approach: good models (diagnostics) would be erroneously overestimated while poor models (diagnostics) would be erroneously underestimated.

Our example of using ECS as a performance metric demonstrates the importance, in the ensemble performance weighting approach, of choosing an appropriate prior distribution for the weight vectors. For any set of weight vectors, the resulting set of the weighted averages (of the diagnostic variable) will have a reduced spread as compared to the unweighted ensemble. How much it is reduced, however, strongly depends on the nature of the set of applied weight vectors. Both, Dirichlet(1) and Dirichlet(1/N), are natural choices for the prior distribution as in both cases the expected value of the weight for each model equals 1/N. With Dirichlet(1/N) large model weights are similarly likely as smaller model weights whereas with Dirichlet(1), on average, the individual model weights tend to be smaller. Therefore, the Dirichlet(1) prior is much more constrained towards values that lie near the expected value than is the Dirichlet(1/N) prior. This, in turn, means that much more data would be necessary, using Dirichlet(1),  to end up with a posterior distribution that assigns most probability mass on values in the tails of the distribution. Thus, we generally recommend to use a Dirichlet prior with shape parameter 1/N and to further make prior predictive checks that ensure that the data that is \emph{a priori} generated by the weighted average models is realistic and largely covers the range of data from the ensemble members.
Moreover, for a reasonable interpretation of the projected likely range of the weighted ensemble, using the expected weight of the posterior distribution obtained from applying the ensemble performance weighting approach, it is necessary to verify how much of the reduced range is effectively attributable to the data and not a mere effect of the weighting itself. Therefore, a comparison between the weighted projections based on the prior weights and based on the posterior weights is a minimal requirement to validate the weighting (see Fig.~\ref{fig:projections-mmm-bma}).

% Paragraph on represented uncertainties
In fact, the interpretation of the projected uncertainty range, and the interpretation of the uncertainties in general, deserve some further discussion. The predicted uncertainty range for our target variable (e.g., future projections of the increase of global mean temperature in Fig.~\ref{fig:projections-mmm-bma}) based on the computed posterior weight vectors in part represents the uncertainty in the computed weights themselves. This is a kind of uncertainty that cannot be included by its nature in the individual performance weighting approach. However, accounting for it does not automatically imply that structural differences between the models are well treated. For large ensembles, this is likely not so problematic, as the ensemble members can be expected to be diverse enough to represent all plausible model trajectories. However, care should be taken when ensembles of opportunity are used, to ensure the uncertainty estimate is not overly narrow.
In addition, to get a more representative estimate of the uncertainty around future projections, it would be important to include other kinds of uncertainties that we have not considered here and are often ignored. This includes the uncertainty of the individual models themselves (which can be estimated based on multiple model runs) as well as the uncertainty of the observations that are used for the performance evaluation. The Bayesian formulation of weighting that we proposed here conveniently allows us to make such additions.

% Caveats to BMA - when is it not appropriate
A limitation of the ensemble performance weighting approach is that it is restricted to linear combinations of individual models. 
However, in some cases it might be more appropriate to combine models in a non-linear way.
Some models, for instance, perform much better in particular regions of the globe than in others, suggesting that the weights may be modeled to differ across grid points, or regions \citep[e.g.,][]{dasbhowmikPerformancebasedMultimodelCombination2021,thaoCombiningGlobalClimate2022}. 
As the choice of the diagnostics, whether or not it is helpful to compute weights dependent on a region or even on a grid-point level also depends on the objective. If the objective is to compute the likely range of global mean temperatures in the future, it seems more plausible to stick with one weight per model whereas weights on the level of grid cells may be useful when considering likely ranges of temperatures within a certain region only.

A key focus of this work has been to compare the different assumptions and implications of the individual  and the ensemble performance weighting approach. 
Overall, past work \citep{massoudGlobalClimateModel2019,massoudBayesianModelAveraging2020} indicate that ensemble performance weighting provides benefits over individual performance weighting. As just mentioned, there may be cases in which ensemble performance weighting is not appropriate. However, there are many cases in the literature where the final goal is to obtain good performance of the combined ensemble, i.e.~we are interested in knowing the weighted-mean prediction. In those cases, ensemble performance weighting is more straightforwardly connected to the goal and provides an estimate of the uncertainty in the weights which is not accounted for in the individual performance weighting approach. If, on the other hand, the goal is simply to rank models or model versions against each other, then the individual performance weighting approach is straightforward in its application. 




